---
title: "Lab 7 Assignment: Group 11"
format:
  html:
    embed-resources: true
editor: visual
---

## GROUP 11:

#### Eleni-Sofia Tseperi s240066, Johanne Lund s233246, Marie Picquet s233736, Eglantine Anton s233242, Wu Qiuyan s241063

## Background

Breast cancer is one of the most prevalent forms of cancer, affecting millions of women worldwide. It occurs when cells in the breast begin to grow uncontrollably, forming a tumour that can be either benign (non-cancerous) or malignant (cancerous). Early detection is critical, as it significantly increases the chances of successful treatment and survival. However, diagnosing breast cancer accurately remains a challenge due to the variability in tumour characteristics.

The data will be working on is the `biopsy` dataset, which comes originally from the MASS package. This breast cancer database was obtained from the University of Wisconsin Hospitals, Madison from Dr. William H. Wolberg. He assessed biopsies of breast tumours for 699 patients up to 15 July 1992; each of nine attributes has been scored on a scale of 1 to 10, and the outcome is also known.

## Aim

## Load Library

```{r}
#| echo: True
#| eval: true
library(tidyverse, verbose = FALSE)
library(ggplot2, verbose = FALSE)

```

## Load Data

```{r}
#| echo: True
#| eval: true
biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

## Data Description

We’ll be analyzing the `biopsy` dataset, which comes originally from the MASS package. It is a breast cancer dataset from the University of Wisconsin Hospitals, Madison from Dr. William H. Wolberg.

He assessed biopsies of breast tumors for 699 patients; each of nine attributes was scored on a scale of 1 to 10. The attributes are: clump_thickness, uniform_cell_size, uniform_cell_shape, marg_adhesion, epithelial_cell_size, bare_nuclei, bland_chromatin, normal_nucleoli, mitoses.

The true outcome (benign/malignant) is also known.

## Analysis
