---
title: "Lab 7 Assignment: Group 11"
format:
  html:
    embed-resources: true
editor: visual
---

## GROUP 11:

#### Eleni-Sofia Tseperi s240066, Johanne Lund s233246, Marie Picquet s233736, Eglantine Anton s233242, Wu Qiuyan s241063

## Background

## Aim

The aim is to use Principal Component Analysis (PCA) to reduce the dimensionality of the biopsy dataset, allowing us to visualize the data in new coordinates (principal components) that capture the most variation, helping to identify patterns and relationships in the breast cancer attributes.

## Load Library

```{r}
#| echo: True
#| eval: true
library(tidyverse, verbose = FALSE)
library(ggplot2, verbose = FALSE)

```

## Load Data

```{r}
#| echo: True
#| eval: true
biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

## Data Description

We’ll be analyzing the `biopsy` dataset, which comes originally from the MASS package. It is a breast cancer dataset from the University of Wisconsin Hospitals, Madison from Dr. William H. Wolberg.

He assessed biopsies of breast tumors for 699 patients; each of nine attributes was scored on a scale of 1 to 10. The attributes are: clump_thickness, uniform_cell_size, uniform_cell_shape, marg_adhesion, epithelial_cell_size, bare_nuclei, bland_chromatin, normal_nucleoli, mitoses.

The true outcome (benign/malignant) is also known.

## Analysis
